---
title: "PCA"
---



```{r}
cc = read.csv("Credit_clients.csv", header=T)
```
```{r}
str(cc)
```
W use the log-transformation for LIMIT_BAL, BILL_AMT1 to BILL_AMT6, PAT_AMT1 to PAY_AMT6 
(log(0) is –Inf, so we add a small value such as 1 to the original data before taking the log). 

```{r}
cc$LIMIT_BAL=log(cc$LIMIT_BAL)
cc[,11:22]=log(cc[,11:22]+1)
```
Next we use the scale( ) function to scale the log-transformed LIMIT_BAL, BILL_AMT1 to BILL_AMT6 and PAT_AMT1 to PAY_AMT6 (for example, scale(LIMIT_BAL)). 

Then we use the function prcomp( ) to run principal component analysis.
```{r}
cc[,2:22]=scale(cc[,2:22])
str(cc)
```
```{r}
cc.pca1=prcomp(cc[,2:22])
cc.pca1
```

We use the plot( ) function to create a scree plot of the principal component variances. 
```{r}
plot(cc.pca1,type="l")
pcadata=as.data.frame(predict(cc.pca1))
eigen(cov(pcadata))
```
As we can see, 4 eigen values are bigger than 1. 

This biplot can be interpreted as follows:
•	Limit_bal and Age are highly correlated and grouped together
•	All the PAY_AMT variables are highly correlated and grouped together
•	All the BILL_AMT variables are highly correlated and grouped together
•	All the Pay variables and Education are highly correlated and grouped together
•	There is no relationship between credit limit and bill amount
•	There is no relationship between payment time and payment amount


Next, we will conduct factor analysis, using the R function factanal( ) to run orthogonal factor analysis with 2, 3 and 4 factors.

```{r}
cc.fa1 = factanal(x=cc,factors=2,score="Bartlett")
cc.fa2 = factanal(x=cc,factors=3,score="Bartlett")
cc.fa3 = factanal(x=cc,factors=4,score="Bartlett")
```
•	According to the results, the 3 factors model is the most interpretable. 
•	Compared to the 2-factor model, the 3-factor model successfully interprets the relationship between Bill_Amt and Pay_AMT. 
•	A 4-factor model seems unnecessary because of its redundancy.

Interpretation:

If the loading value of one variable is more than 0.5, this variable is positively correlated with the factor. 
On the other hand, if the loading value of one variable is smaller than -0.5, this variable is negatively correlated with the factor. 
Factor 1 and Factor 3 are highly correlated.

To plot the results of factor analysis, we will need to create a function.
First we will plot the two-factor model
```{r}
biplot.fa =  function (fa,...) {
  x = fa$scores[,1:2]
  y = fa$loadings[,1:2]
  biplot(x,y,...)
}
```
```{r}
biplot.fa(cc.fa1,cex=c(0.2,1.2))
```
Interpretation:
•	Limit_bal and Age are highly correlated and grouped together
•	All the PAY_AMT variables are highly correlated and grouped together
•	All the BILL_AMT variables are highly correlated and grouped together
•	All the Pay variables and Education are highly correlated and grouped together
•	There is no relationship between credit limit and bill amount
•	There is no relationship between payment time and payment amount

The results are similar to the PCA model.


Next, we will try the oblimin rotation for oblique EFA. For this we install the package "GPArotation" for the oblimin rotation, and create a 3-factor EFA model using the rotation oblimin.

```{r}
install.packages("GPArotation")
```
```{r}
library(GPArotation)
cc.fa4 = factanal(cc,factors = 3,rotation = "oblimin", scores = "Bartlett")
cc.fa4
```
In general, the loading values are same in the 3-factor model without rotation. 

But this rotation method increases the performance a little more because the loading values increase.

